---
title: "Technological Mediation"
subtitle: "A Postphenomenology Primer for Instructors, Designers, and More"
author: "Dr. Ryan Straight"
date: "2021 September 21"
output:
  xaringan::moon_reader:
    css: ["assets/arizona-fonts.css", "assets/arizona.css", "default"]
    self_contained: false # if true, fonts will be stored locally
    seal: true # show a title slide with YAML information
    nature:
      beforeInit: ["assets/remark-zoom.js"]
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9' # alternatives '16:9' or '4:3' or others e.g. 13:9
      navigation:
        scroll: false # disable slide transitions by scrolling
---
```{r setup, include=FALSE}
# A University of Arizona theme for Xaringan RMarkdown slide decks

# Based on [Garth Tarr's `sydney_xaringan` theme](https://github.com/garthtarr/sydney_xaringan), itself based on the original [xaringan](https://github.com/yihui/xaringan) theme by Yihui Xie.

# Fonts, colors, and other assets from [brand.arizona.edu](https://brand.arizona.edu). Please note that _Proximia Nova_ is a licensed font that should only be used with permission, so is not included here. Instead, MiloWeb is used.

options(htmltools.dir.version = FALSE)
library(xaringan)
library(xaringanExtra)
library(DT)
library(kableExtra)
library(tidyverse)

# Xaringan extras
xaringanExtra::use_scribble()     #<< turns on Scribbling

# Code highlighting on over
xaringanExtra::use_extra_styles(
  hover_code_line = TRUE,         #<<
  mute_unhighlighted_code = TRUE  #<<
)
# xaringanExtra::use_slide_tone() # unmute when publishing

# Turns on clipboard options. Great for giving code to students.
htmltools::tagList(               # # turns on the clipboard option
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa fa-clipboard\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
    error_text = "<i class=\"fa fa-times-circle\" style=\"color: #F94144\"></i>"
  ),
  rmarkdown::html_dependency_font_awesome()
)

# Editable slides
xaringanExtra::use_editable(expires = 1)
```

```{r, load_refs, echo=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
library(RefManageR)
BibOptions(check.entries = FALSE, 
           bib.style = "authoryear", 
           cite.style = 'authoryear', 
           style = "markdown",
           hyperlink = FALSE, 
           dashed = FALSE)
myBib <- ReadBib("assets/example.bib", check = FALSE)
```

# Today's schedule

.pull-left[

+ Intros
+ What is postphenomenology?
+ Why is it relevant to us?
+ How do you *do* it?

## Questions during?

## Ask away!

]

.pull-right[

<div style="width:100%;height:0;padding-bottom:100%;position:relative;"><iframe src="https://giphy.com/embed/ZbOXZEugwT26awakGe" width="100%" height="100%" style="position:absolute" frameBorder="0" class="giphy-embed" allowFullScreen></iframe></div><p><a href="https://giphy.com/gifs/latenightseth-lol-seth-meyers-lnsm-ZbOXZEugwT26awakGe">via GIPHY</a></p>

]

---
class: secondary-silver
background-image: url("assets/ua_horiz_rgb_black_4.svg")
background-size: 20%
background-position: 95% 95%

# Introductions

(icon)

---

# Dr Ryan Straight

(slide about me)

---
class: secondary-river
background-image: url("assets/ua_horiz_rgb_black_4.svg")
background-size: 20%
background-position: 95% 95%

# What is postphenomenology?

(icon)

???

I'll be giving a brief overview of what PP is and where it comes from, but the majority of the work in terms of "doing" PP and its integration into this online learning space actually comes from the work of Catherine Adams and others. So, using her definition:

>"phenomenology that attends to specific technologies and the existential and epistemological differences they may be making to the lifeworld" (2018, p.4).

And we'll unpack that as this session goes on. First, how did we get from traditionally phenomenology to ... this?

---
class: brand-blue

# A brief history of postphenomenology


...

???

+ Ihde in the 70s
+ Ihde in the 90s
+ Verbeek and others
+ Actor Network Theory
+ Technological mediation
+ Applied postphenomnology, or Adams' "postphenomenology of practice," as her specific version is based on the phenomenological work of Van Manen.

So let's start off really digging into this framework with some terminology. Listen closely as there will be a quiz.

---
class: brand-blue

# Terminology


---
class: brand-red

## Technological mediation

> Technology *necessarily* mediates our experience with the world.

???

When we think of how we interact with the world, in classical phenomenology we're encouraged (required, in fact) to "block" or otherwise attempt to eschew all elements of the lifeworld that can interfere with our experience to get to the root of the thing. PP says, no, technology (as we experience it now) shouldn't be dismissed or considered a hindrance but rather something to be empirically studied. And we'll look at just how postphenomenologists like Ihde typically envision this in just a little bit.

--

## Multistability

> Things have more than one use, technology especially.

???

Think of a hammer. It can be a construction tool, a paperweight, or a weapon. Don't confuse this with *convergence*, for example with a phone now being a GPS, a camera, and so on, as these various technologies have converged into one. No, here we're talking about a technology that is built for one thing but has various (multiple) different possibly unintended functions. 

--

## Transparency

> When technology fades from "view."

???

If you recall this from your philosophy of technology days, the hammer is a tool but becomes something else entirely when it breaks. When it's functioning, though, a sense of embodiment arises and the hammer and hammer user become one. Like an air conditioner: you don't really think much about it until it doesn't work. If you're a gamer, think of this as the controller: eventually you stop feeling that technology as something opaque between you and the screen. 

And this leads us to the fun stuff: technic relations.

---
class: secondary-silver

# Technic Relations

(icon)

???

The real goal of all this--and, hopefully, why you're here--is to discuss just how we relate to the world through, because of, and maybe even sometimes *in spite of* technologies.


---
class: brand-blue

# Embodiment

...


---
class: brand-blue

# Hermenutic

...


---
class: brand-blue

# Alterity

...


---
class: brand-blue

# Background

...

???

The air conditioner we talked about. 

Now, time for a little thought experiment. I want you to imagine a typical online student, however they may look.

---
class: secondary-river

# Thought experiment

(icon)

???

Do you have it in your mind? Picture everything. The chair, the desk, the laptop or PC, the ambient sounds, the temperature, the lighting. Now picture what kind of LMS is coming through. Is it responsive? Think about the PDF the student is reading on the screen: is it OCR friendly? 
Think about all these things and start categorizing them into the different relations: the mouse and keyboard are part of the student, so the embodiment relation is occurring. The way the course is designed is unfamiliar, so there's some breakdown in the hermeutic relationship as the student tries to interpret what's presented. 

And so on.

And what about when the student is in a synchronous online class, like via Zoom?

Zoombie?

---









































<!-- Ending slide below -->
---
class: arizona-blue
background-image: url(assets/ua_horiz_rgb_4.svg)
background-size: 260px
background-position: 5% 95%

# Thank you!

Contact info:


.content-box-bloom[You can find references and readings on the [MAVRX Lab's event page](https://mavrxlab.org/event/ac21/)!]

.pull-right[.pull-down[

ryanstraight@arizona.edu

<br><br><br>

]]
